---
title: "The Validation Set Approach"
author: "Lisa Stuart"
date: "March 17, 2016"
output: html_document
---


1. Set a seed for Reproducible Research.
2. Split the n observations into two sets of approximately equal size, a training set and a validation set.
a. Fit the model using the training observations.  Let beta_hat denote the regression coefficient estimates.
b. For each observation in the validation set, compute the training error
        e_i = (y - y_hat)^2
3. Calculate the total validation set error by summing the e_i's over all of the validation set observations.

### An Example from Introduction to Statistical Learning in R

```{r example}
library(ISLR)
set.seed(1) 
dim(Auto)
```

Randomly sample 196 numbers out of first 392 integers to create training set. Printing the training set shows the indices the training set has:

```{r sampling}
train=sample(392,196)
train
```

Fit a linear model to the training set and then calculate the test error from it; y = mpg, predict = y_hats, all but train subset gives test subset

```{r linear model fit}
attach(Auto)
lm.fit=lm(mpg~horsepower,data=Auto,subset=train)
mean((mpg-predict(lm.fit,Auto))[-train]^2)
```

Fit a 2nd degree polynomial (quadratic) 

```{r 2nd degree}
lm.fit2=lm(mpg~poly(horsepower,2),data=Auto,subset=train)
mean((mpg-predict(lm.fit2,Auto))[-train]^2)
```

Fit a 3rd degree polynomial

```{r 3rd degree}
lm.fit3=lm(mpg~poly(horsepower,3),data=Auto,subset=train)
mean((mpg-predict(lm.fit3,Auto))[-train]^2)
```

Fit a 5th degree polynomial, a more complex model but not lower test error

```{r 5th degree}
lm.fit5=lm(mpg~poly(horsepower,5),data=Auto,subset=train)
mean((mpg-predict(lm.fit5,Auto))[-train]^2)
```

Check training error of 3rd degree polynomial:

```{r training error}
mean((mpg-predict(lm.fit3,Auto))[train]^2)
```

Repeat with a different seed:

```{r repeat}
set.seed(2)
train=sample(392,196)
lm.fit=lm(mpg~horsepower,subset=train)
mean((mpg-predict(lm.fit,Auto))[-train]^2)
lm.fit2=lm(mpg~poly(horsepower,2),data=Auto,subset=train)
mean((mpg-predict(lm.fit2,Auto))[-train]^2)
lm.fit3=lm(mpg~poly(horsepower,3),data=Auto,subset=train)
mean((mpg-predict(lm.fit3,Auto))[-train]^2)
```

